Quantum Hoare Types

### 2019-11-07 paper/paper.tex
Add partial teleportation example
Add explanation of states for teleport example

	\item Entanglement Swapping (variant of teleportation):\\
	Initially Alice ($a_1$) and Carol ($c$) share an EPR pair and so do Alice ($a_2$) and Bob ($b$)
\[ \begin{array}{rcl}
\hoare{(c, a_1), (a_2, b): \E}{CX(a_1, a_2);}{??} \\
\hoare{??}{H(a_1)}{??}\\
\hoare{??}{meas(a_1); \; meas(a_2);}{(c,b): \E; a_1, a_2 : qbit}\\
\end{array} \]

@@

@@

>>>>

\newpage

	\item Quantum Teleportation

	We have three qubits. Qubit 0 represents the message in state $\ket{\psi} = \alpha \ket{0} + \beta \ket{1}$. Qubit 1 corresponds to Alice and qubit 2 corresponds to Bob and they both start in the state $\ket{0}$. To send the message, first step in this protocol is for qubits 1 and 2 to be entangled.
\[ \begin{array}{rcl}
\hoare{q_0: qbit = \ket{\psi}; q_1, q_2 : \Z}{H(q_1)}{q_0: qbit; q_2 : \Z; q_1 : \X} \\
\hoare{q_0: qbit = \ket{\psi}; q_2 : \Z; q_1 : \X}{CX(q_1, q_2)}{q_0: qbit; (q_1,q_2) : \E} \\
\end{array} \]

Once entangled they are in the state $\ket{\phi^+} = \frac{\ket{0} + \ket{1}}{2}$. The complete quantum state looks like:

\[  \ket{\psi}\ket{\phi^+} = (\alpha\ket{0} + \beta\ket{1})(\frac{1}{\sqrt{2}}(\ket{00} + \ket{11})) \]

or, normalized:

\[  \ket{\psi}\ket{\phi^+} = \frac{\alpha}{\sqrt{2}}\ket{000} + \frac{\alpha}{\sqrt{2}}\ket{011} + \frac{\beta}{\sqrt{2}}\ket{100} + \frac{\beta}{\sqrt{2}}\ket{111} \]

In the second step we apply CNOT to qubits 0 and 1 and then a Hadamard to qubit 0.

\[ \begin{array}{rcl}
\hoare{q_0: qbit = \ket{\psi}; (q_1,q_2) : \E}{CX(q_0, q_1)}{??} \\
\hoare{??}{H(q_0)}{??} \\
\end{array} \]

After the CNOT application, the resulting state is:

\[ \frac{\alpha}{\sqrt{2}}\ket{000} + \frac{\alpha}{\sqrt{2}}\ket{011} + \frac{\beta}{\sqrt{2}}\ket{110} + \frac{\beta}{\sqrt{2}}\ket{101} \]

and after Hadamard application to qubit 0, it becomes:

\[ \frac{\alpha}{\sqrt{2}}(\frac{1}{\sqrt{2}}(\ket{0} + \ket{1}))\ket{00} + \frac{\alpha}{\sqrt{2}}(\frac{1}{\sqrt{2}}(\ket{0} + \ket{1}))\ket{11} + \frac{\beta}{\sqrt{2}}(\frac{1}{\sqrt{2}}(\ket{0} - \ket{1}))\ket{10} + \frac{\beta}{\sqrt{2}}(\frac{1}{\sqrt{2}}(\ket{0} - \ket{1}))\ket{01} \]

This can be simplified to:

\[ \frac{1}{2}\big[\alpha\ket{00\;0} + \beta\ket{00\;1} + \alpha\ket{01\;1} + \beta\ket{01\;0} + \alpha\ket{10\;0} - \beta\ket{10\;1} + \alpha\ket{11\;1} - \beta\ket{11\;0}\big] \]

The next step is the measurement of the first two qubits:

\[ \begin{array}{rcl}
\hoare{??}{meas(q_0); \; meas(q_1);}{q_0, q_1 : bit, q_2: qbit}\\
\end{array} \]

This leads to 4 possibilities of getting $(0, 0), (0, 1), (1, 0, (1, 1)$. Depending on what we got, we need to make a correction on the resulting state of qubit 2, so we get the original message $\ket{\psi}$. This is done by applying X and Z gates conditioned on the two measured bits.

\[ \begin{array}{rcl}
\hoare{q_0, q_1 : bit, q_2: qbit}{CX(q_1, q_2); \; CZ(q_0, q_2);}{q_0, q_1 : bit, q_2: qbit = \ket{\psi}}\\
\end{array} \]

\TODO{Need a way to assert that $q_2$ has the same quantum state at the end of the program as $q_0$ did at the beginning}

	\item Entanglement Swapping (variant of teleportation):\\
	Initially Alice ($a_1$) and Carol ($c$) share an EPR pair and so do Alice ($a_2$) and Bob ($b$)
\[ \begin{array}{rcl}
\hoare{(c, a_1), (a_2, b): \E}{CX(a_1, a_2);}{??} \\
\hoare{??}{H(a_1)}{??}\\
\hoare{??}{meas(a_1); \; meas(a_2);}{(c,b): \E; a_1, a_2 : bit}\\
\end{array} \]

@@

\TODO{Can we assign such type refinements to all/most quantum state patterns that have recognizable structure for human programmers?}

@@

\listofnotes
